dbRDA analysis

### 2017-04-14 doc/RMD/RDA-analysis.Rmd
update RDA analysis (I think?)

```{r, load_packages}
library('tidyverse')
library('poppr')
library('vegan')
```

```{r, load_data}
load("data/sclerotinia_16_loci.rda")
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
setPop(dat) <- ~Host/Source/Region/Year
dat11 <- dat[loc = keeploci, mlg.reset = TRUE]
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat16cc   <- clonecorrect(dat, ~Host/Source/Region/Year, keep = 1:4)
dat11cc
dat16cc
# environmental variables for dbRDA
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
ENV11 <- bind_cols(select(other(dat11cc)$meta, Severity), 
                   select(strata(dat11cc), -MCG))
ENV11
stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))
ENV16 <- bind_cols(select(other(dat16cc)$meta, Severity), 
                   select(strata(dat16cc), -MCG))
ENV16
```

## Calculate Bruvo's distance

```{r}
bruvo11cc <- bruvo.dist(dat11cc, replen = other(dat)$REPLEN)
bruvo16cc <- bruvo.dist(dat16cc, replen = other(dat)$REPLEN)
```

```{r}
commtab <- mlg.table(dat11cc, plot = FALSE)
cap11cc <- capscale(bruvo11cc ~ Region + Host + Year + Severity, 
                    data = ENV11, 
                    comm = , 
                    add = TRUE)

```

```{r}
capsum <- summary(cap11cc)
summary(summary(cap11cc))
##########Extracting 10 most important centroids for factor constrains 
###############It doesn't matter how you start the ggplot aesthetics. Can be DecendCentro CAP1 and CAP2 or site_centroids
############### CAP1 and CAP2. Same graph is resulted

Centroids <- data.frame(capsum$centroids) #no need to convert to a dataframe
LenCent <- sqrt((Centroids[,1]^2)*(Centroids[,2]^2))   ###  This calculates the arrow length (ie. strength of the eigenvector)
NewCentro <- cbind(Centroids, LenCent)
DecendCentro <- head(NewCentro[order(NewCentro[,7], decreasing = T),], n=23)   ###  this finds the n strongest eigenvectors by ordering them
#above "head" is correct

DecendCentro$cent_type <- rownames(head(DecendCentro, n=23))
DecendCentro <- head(DecendCentro, n = 10)
DecendCentro #constraint factor cetroids

## Testing plot in ggplot

ggplot(data= DecendCentro, mapping = aes(x = CAP1, y = CAP2))+
  geom_text(aes(label = cent_type))+ # label tells geom_text which text you want to plot
  ylim(-3.5,3.5)+ #NB note that this is a convenience wrapper and may cut data out of your plot
  xlim(-3,3.5) #important if you are calculating stats in the plot - these will be excluded

# sites/isolates
site_centroids <- data.frame(capsum$sites)
site_centroids$isolate_names <- rownames(site_centroids) # you can also put label = rownames() etc in the ggplot code. But sometimes you may want it in a column for later

##plotting both main centroids and isolates together
# I am plotting isolates firest and then envt centroids, alpha = transparancy
ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
  geom_point(data = site_centroids, colour = "orange", size =3.5, alpha =0.7)+
  geom_text(aes(label = cent_type))+
  ylim(-4,3.5)+
  xlim(-3.5,3.5)

#######strongest arrows: THESE ARE BIPLOT scores#############

arrws <- data.frame(capsum$biplot)
lenarrw <- sqrt(arrws[,1]^2*arrws[,2]^2)   ###  This calculates the arrow length (ie. strength of the eigenvector)
arrws <- cbind(arrws, lenarrw)
arrws <- head(arrws[order(arrws[,7], decreasing = T),], n=10)   ###  this finds the n strongest eigenvectors by ordering them
arrws

arrws$class <- (rownames(arrws)) #turning rownames into a variable
mult <- attributes(capsum)$const/2.5 # 97% sure this is the scaling for the arrows, pls let me know if I'm wrong
arrws$class
#arrws$class <- c("MX", "ND", "CA", "mtc", "NE", "MCG ")  # change this if you want to change how names are displayed
#arrws$class <- c("MX", "mtc", "NE", "ND", "MCG", "CA")
##locations abbreviations; see the original for full names

#####plotting centroids, isolates and biplot arrows together#########
##@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@########

arrws$class <- c("Mexico", "2005", "CA", "Vista", "CO", "Host: unknown", "OR", "Severity", "MN", "WA")
mult <- attributes(capsum)$const/3

(Cap_plot <- ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
   geom_text(aes(label = cent_type), colour = "blue")+
   geom_point(data = site_centroids, alpha=1/2, colour = "dark orange", size = 2.5, )+
   #coord_cartesian(x = c(-3.5, 3.5), y = c(-4.5, 2.5))+# For bruvo
   coord_cartesian(x = c(-3.5, 3.5), y = c(-2.5, 5))+
   geom_segment(data = arrws,
                aes(x = 0, xend = mult * CAP1, y = 0, yend = mult * CAP2),
                arrow = arrow(length = unit(0.3, "cm")), colour = "dark grey") + #grid is required for arrow to work.
   geom_text(data = arrws, color = "black",
             aes(x= (mult + mult/10) * CAP1, y = (mult + mult/10) * CAP2, label = arrws$class), #otherwise you could use hjust and vjust. I prefer this option
             size = 5, hjust = 0.5)+ theme_bw())
```

>>>>

```{r, load_packages}
library('tidyverse')
library('poppr')
library('vegan')
library('ggrepel')
```

```{r, load_data}
load("data/sclerotinia_16_loci.rda")
setPop(dat)   <- ~Host/Source/Region/Year
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat16cc <- clonecorrect(dat, ~Host/Source/Region/Year, keep = 1:4)
dat11cc
dat16cc
# environmental variables for dbRDA
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))

makeENV <- function(DAT, CC){
  # Creating the data frame with severity
  META   <- select(other(DAT)$meta, -Isolate)
  STRATA <- select(strata(DAT), -MCG)
  MLL    <- data.frame(MLG = mll(DAT))
  sev    <- bind_cols(META, STRATA, MLL) %>%
    group_by(Host, Source, Year, Region, MLG) %>% 
    summarize(Severity = mean(Severity)) %>% # Get mean severity per MLG
    ungroup()
  # Ensuring the data is in the correct order
  META   <- select(other(CC)$meta, -Isolate)
  STRATA <- select(strata(CC), -MCG)
  MLL    <- data.frame(MLG = mll(CC))
  bind_cols(META, STRATA, MLL) %>% 
    left_join(sev)
}

ENV11 <- makeENV(dat11, dat11cc)
ENV11
stopifnot(identical(ENV11$MLG, mll(dat11cc)))
ENV11 <- select(ENV11, -MLG)
ENV16 <- makeENV(dat, dat16cc)
ENV16
stopifnot(identical(ENV11$MLG, mll(dat11cc)))
ENV16 <- select(ENV16, -MLG)
```

## Function to tie everything together

```{r}
choose_dbrda <- function(bdist, ENV, CHOOSER = "ordistep", ...){
  mod0  <- capscale(bdist ~ 1, data = ENV, add = TRUE)
  mod1  <- capscale(bdist ~ ., data = ENV, add = TRUE)
  CHOOSER <- match.fun(CHOOSER)
  the_model <- CHOOSER(mod0, scope = formula(mod1), ...)
  return(the_model)
}
arrowMul <- function(arrows, data, at = c(0, 0), fill = 0.75) {
    u <- c(range(data[,1], range(data[,2])))
    u <- u - rep(at, each = 2)
    r <- c(range(arrows[, 1], na.rm = TRUE), range(arrows[, 2], na.rm = TRUE))
    rev <- sign(diff(u))[-2]
    if (rev[1] < 0)
        u[1:2] <- u[2:1]
    if (rev[2] < 0)
        u[3:4] <- u[4:3]
    u <- u/r
    u <- u[is.finite(u) & u > 0]
    fill * min(u)
}
plot_dbrda <- function(db){
  dbsum     <- scores(db, display = c("cn", "bp", "sites"), scaling = "sites")
  Centroids <- as.data.frame(dbsum$centroids)
  Centroids <- rownames_to_column(Centroids, var = "cent_type")
  Centroids <- mutate_(Centroids, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
  # Centroids
  SampleCentroids <- rownames_to_column(data.frame(dbsum$sites), var = "isolate_names")

  mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites) * 10
  Arrows  <- data.frame(dbsum$biplot * mul)
  Arrows  <- rownames_to_column(Arrows, var = "class")
  Arrows  <- mutate_(Arrows, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
  Arrows  <- arrange(Arrows, Length)
  Arrows  <- top_n(Arrows, 8)
  ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
   geom_point(data = site_centroids, alpha = 1/2, fill = "dark orange", color = "black", size = 2.5, pch = 21)+
   coord_cartesian() +
   geom_segment(aes(x = 0, xend = CAP1, 
                    y = 0, yend = CAP2),
                arrow = arrow(length = unit(0.3, "cm")), 
                data = Arrows
                ) + 
   geom_label_repel(aes(x = CAP1, y = CAP2, label = class), data = Arrows)
}
```

# Calculations

```{r rda_calculation, cache = TRUE}
dat11cc.bruvo <- dat11cc %>% bruvo.dist(replen = other(.)$REPLEN)
cap11cc       <- choose_dbrda(dat11cc.bruvo, ENV = ENV11, CHOOSER = "ordistep")
dat16cc.bruvo <- dat16cc %>% bruvo.dist(replen = other(.)$REPLEN)
cap16cc       <- choose_dbrda(dat16cc.bruvo, ENV = ENV16, CHOOSER = "ordistep")
```

```{r}
plot_dbrda(cap11cc) + theme_bw() + theme(text = element_text(size = 14))
plot_dbrda(cap16cc) + theme_bw() + theme(text = element_text(size = 14))
```
